---
title: "Selecting locations for NE Greenlip Timed-swim surveys"
author: "Jaime McAllister and Craig Mundy"
date: "`r Sys.Date()`"
output:
  pdf_document:
    fig_height: 8
    fig_width: 8
    # includes:
    #   in_header: header.tex
    toc: yes
  word_document:
    fig_caption: yes
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = "",     ## nothing goes in front of each line
                      message = FALSE,  ## into the console, not the document
                      warning = FALSE,  ## into the console, not the document
                      fig.align = "center",
                      fig.show = "hold",
                      out.height = "0.8\\textwidth",
                      out.width = "0.8\\textwidth")
options(knitr.kable.NA = '')


suppressPackageStartupMessages({
library(tictoc)
library(sf)
library(nngeo)
# library(maptools)
# library(rgdal)
library(sp)
library(spatialEco)  
# library(rgeos)
library(lubridate)
# library(spdplyr)
library(tidyverse)
library(tmap)
library(tictoc)
library(spdep)
library(openxlsx)  
})

## set EPSG CRS

GDA2020 <- st_crs(7855)
GDA94 <- st_crs(28355)
WGS84 <- st_crs(4326)


```


# Read in hex layer
Read in grid wide spatial layer, and transform to GDA2020. This hex cell layer contains the majority of fishing activity in the closed blocks.
```{r prepare base layers}
## Read in greenlip abalone wide hex layer

ab.hex <- readRDS(paste(sprintf("C:/Users/%s/Dropbox (UTAS Research)/DiveFisheries/GIS/SpatialLayers/TimeSwimLayers/grid1HA_greenlip.rds", Sys.info()[["user"]])))


## convert to sf
sf.ab.hex <- st_as_sf(ab.hex)

## Transform from GDA94 to GDA2020
sf.ab.hex <- st_transform(sf.ab.hex, GDA2020)
```

# Add quartile to dataframe
THe input variable is the total catch in Kg for each cell from 2012 - 2019. We use the dplyr ntile() function to calculate five quantiles (0-20, 20-40, etc). based on total catch.
```{r add quartile var}

# Filter hex layer to required block and add quartile
ts.hex <- sf.ab.hex %>% 
  mutate(subblockno = gsub(" ", "", subblockno)) %>% 
  filter(subblockno %in%  c('31A', '31B', '39A', '39B')) %>% 
  within({
    cell.ntile <- ntile(GLhexcatch, 5)
  })

outname.ts.hex <- sprintf("C:/Users/%s/Dropbox (UTAS Research)/DiveFisheries/GIS/SpatialLayers/TimeSwimLayers/grid1HA_greenlip.gpkg", Sys.info()[["user"]])
st_write(ts.hex, dsn = outname.ts.hex, layer = "pointqt", driver = "GPKG", append = F)

hex.cell.summary <- ts.hex %>%
  st_set_geometry(NULL) %>%
  group_by(cell.ntile) %>%
  summarise(
    mncatch = mean(GLhexcatch, na.rm = T),
    catch = sum(GLhexcatch, na.rm = T),
    n = n()
  ) %>%
  print()

```

# randomly sample cells within strata.
We want to randomly select 150 hex cells across the top three quantile strata. The two lower quantile strata contribute `r sum(hex.cell.summary$catch[1:2])/sum(hex.cell.summary$catch)`  of a total catch of `r sum(hex.cell.summary$catch) `. The top three quantile groups contribute `r sum(hex.cell.summary$catch[3])/sum(hex.cell.summary$catch)`, `r sum(hex.cell.summary$catch[4])/sum(hex.cell.summary$catch)`  and `r sum(hex.cell.summary$catch[5])/sum(hex.cell.summary$catch)`  of the catch, respectively.
```{r random sample of cells}

# This code was originally used for site selection in the 2023-2024 timed swim validation research experiment for the NE greenlip and Actaeons.

set.seed(123) 

# set sample size required (i.e. n = 150)
samp.size <- 150

# Exclude oid where a gps was left on while transiting on a motherboat.

oid_exc <- c(1062044, 1062045, 1062604, 1062605, 1062606, 1062607, 1062608, 1062609, 1062610, 1062611, 1062614, 1062632, 1062633, 1062634, 1062635, 1063208, 1063209, 1063210, 1063211, 1063212, 1063213, 1063214, 1063215, 1063216, 1063217, 1063218, 1063219, 1063220, 1063221, 1063222, 1063223, 1063224, 1063225, 1063226, 1063227, 1063228)

# randomly select sites (NOTE: manually change dataframe name for each block)
cellqt_NE_GL <- ts.hex %>% 
  filter(!oid %in% oid_exc &
           cell.ntile > 2) %>%
  group_by(blockno, cell.ntile)  %>% 
  st_as_sf() %>%
  subsample.distance(size = samp.size, d = 50) %>% 
  st_as_sf() %>% 
  ungroup()

  # quick summary of sites x block x strata
table(cellqt_NE_GL$blockno, cellqt_NE_GL$cell.ntile)

```
# Examine Nearest Neighbour distances
We find the  k=1 Nearest Neighbour and then calculate the distance between a cell and its closest neighbor. Centroids of hex cells are ~ 107m apart.

```{r examine nnb}
knear <- knearneigh(st_centroid(cellqt_NE_GL), k = 1, longlat = FALSE, use_kd_tree = TRUE)
neighbors <- knn2nb(knear)

summary.nb(neighbors, st_coordinates(st_centroid(cellqt_NE_GL)), longlat = NULL, zero.policy=TRUE)

coords.hex <- st_coordinates(st_centroid(cellqt_NE_GL))
nc_sp <- as(cellqt_NE_GL, 'Spatial')
cellqt_NE_GL.Knbdist <- nbdists(neighbors, st_coordinates(st_centroid(cellqt_NE_GL)), longlat = NULL)

nbdists <- unlist(cellqt_NE_GL.Knbdist) %>% as.tibble()

nbdists %>% filter(nbdists < 125) %>% count()

colnames(nbdists) <- "nbDist"

nbdists %>% 
  ggplot(aes(nbDist)) +
  geom_histogram()

nbdists %>% 
  filter(nbDist < 2000) %>% 
  ggplot(aes(nbDist)) +
  geom_histogram()
```

# Plot of nnb connections
```{r examine nnb network}

neighbors_sf <- as(nb2lines(neighbors, coords = st_coordinates(st_centroid(cellqt_NE_GL))), 'sf' )
neighbors_sf <- st_set_crs(neighbors_sf, st_crs(cellqt_NE_GL))

ggplot(cellqt_NE_GL) + 
  geom_sf(fill = 'salmon', color = 'white') +
  geom_sf(data = neighbors_sf) +
  geom_sf(data = st_centroid(cellqt_NE_GL)) +
  theme_minimal() +
  ylab("Latitude") +
  xlab("Longitude")



```


# Export files
It is much easier to explore the data in QGIS. We export the selected cells as a hex layer and a point layer (centroid of the hex).

```{r export to gpkg}

# These are the original sites chosen for the 2023-2024 timed swim validation research experiment for the NE greenlip and Actaeons.

## Export cells to gpkg ####
pointqt <- st_centroid(cellqt_NE_GL)
outname.point <- sprintf("C:/Users/%s/University of Tasmania/IMAS-DiveFisheries - Assessments - Documents/Assessments/GIS/SpatialLayers/TimeSwimLayers/TimedSwim_NE_GL_2023_50m.gpkg", Sys.info()[["user"]])
st_write(pointqt, dsn = outname.point, layer = "pointqt", driver = "GPKG", append = F)

# Convert geometry to latitude and longitude to create dataframe
pointqt_df <- pointqt %>%
        st_transform(crs = st_crs(4326)) %>%
  sfheaders::sf_to_df(fill = T) %>%
  select(c(zone, blockno, subblockno, cell.ntile, oid, x, y)) %>%
  dplyr::rename('latitude' = y,
                'longitude' = x) %>% 
 arrange((oid)) %>% 
 mutate(site_order = row_number(),
        site = paste('GL', 2023, blockno, site_order, sep = '-'))

pointqt_sf  <-  pointqt_df %>% 
  st_as_sf(coords = c('longitude', 'latitude'), crs = st_crs(4326))
 
# Export Excel file 
write.xlsx(pointqt_df, sprintf("C:/Users/%s/University of Tasmania/IMAS-DiveFisheries - Assessments - Documents/Assessments/GIS/SpatialLayers/TimeSwimLayers/TimedSwim_NE_GL_2023_50m.xlsx", Sys.info()[["user"]]), 
           sheetName = "Sheet1",
           col.names = TRUE, row.names = TRUE, append = FALSE)
 
# Export Excel file ready for GPX external file creation for plotter
# GDAL package for creating GPX files no longer supported in R
# pointqt_gpx <- pointqt_df %>% 
#    select(site, longitude, latitude) %>% 
#    rename('waypointid' = site)

pointqt_gpx <- pointqt_df %>% 
 dplyr::rename('name' = 'site') %>% 
 select(c(name, longitude, latitude)) %>% 
 add_column(description = NA, 
            comment = NA)

write.xlsx(pointqt_gpx, sprintf("C:/Users/%s/University of Tasmania/IMAS-DiveFisheries - Assessments - Documents/Assessments/GIS/SpatialLayers/TimeSwimLayers/TimedSwim_NE_GL_2023_GPX.xlsx", Sys.info()[["user"]]), 
           sheetName = "Sheet1",
           col.names = TRUE, row.names = F, append = FALSE)
 
```
```{r reference site selection 2025}

# In 2025, following the successful trial of the timed swim methodology in the validation research experiment, it was decided to continue monitoring the status of the NE greenlip fishery. Data from the 2023–2024 experiments would serve as the baseline abundance data. To maintain consistency with the methodology used in East Coast blacklip timed swim surveys, a random selection of 15 reference sites was chosen from these baseline sites. A 50 m buffer was then applied to these sites, and 45 new sites were randomly selected from the remaining locations outside this buffer for surveys in subsequent years.

# Read GL layer of sites surveyed in 2023/24 timed swim validation survey
gl_sites_surveyed <- st_read(paste(sprintf("C:/Users/%s/Dropbox (UTAS Research)/DiveFisheries/GIS/SpatialLayers/TimeSwimLayers/TimedSwim_NE_GL_2024_sites_sampled_2024-10-24.gpkg", Sys.info()[["user"]])))

# Read in original sites selected for the 2023/24 survey
gl_sites_proposed <- read.xlsx(paste(sprintf("C:/Users/%s/Dropbox (UTAS Research)/DiveFisheries/GIS/SpatialLayers/TimeSwimLayers/TimedSwim_NE_GL_2023_50m.xlsx", Sys.info()[["user"]])), detectDates = T) %>% 
 select(-1)

# Join sites sureyed with their original OID and cell.ntile value
gl_sites <- left_join(gl_sites_surveyed, gl_sites_proposed %>% select(site, oid, cell.ntile, blockno, subblockno))

# Set seed
set.seed(123)

# Randomly select 15 reference sites from survey data
gl_ref_sites_sf <- gl_sites %>%
 distinct(site, .keep_all = T) %>% 
 sample_n(15) %>% 
 mutate(ref_site = 'ref')

# Unset seed
set.seed(NULL)

# Create dataframe of reference sites
gl_ref_sites_df <- gl_ref_sites_sf %>%
 st_as_sf(coords = c("x", "y")) %>% 
  st_set_crs(st_crs(7855)) %>%
  st_transform(crs = st_crs(4326)) %>%
  sfheaders::sf_to_df(fill = T) %>%
  select(site, blockno, subblockno, oid, cell.ntile, x, y, ref_site) %>% 
  dplyr::rename('latitude' = y,
                'longitude' = x) %>% 
  mutate(blockno = as.integer(blockno),
         cell_ntile = as.integer(cell.ntile)) %>% 
 select(-cell.ntile)

# Export Excel version of reference sites to survey year folder
data_folder <- file.path(paste(sprintf('C:/Users/%s/Dropbox (UTAS Research)/DiveFisheries/Abalone/FISdata',
                                            Sys.info()[["user"]])), paste('FIS_TimedSwimSurveys2025', sep = ''))

outname_point <- paste(data_folder, paste('/TimedSwimSites_NEGreenlip_ReferenceSites_Final_2025', ".gpkg", sep = ""), sep = '')
st_write(gl_ref_sites_sf, dsn = outname_point, layer = "pointqt", driver = "GPKG", append = F)

gl_ref_sites_df %>%
 write.xlsx(file = paste(data_folder, paste('/TimedSwimSites_NEGreenlip_ReferenceSites_Final_2025.xlsx', sep = ''), sep = ''), sheetName = "Sheet1", col.names = TRUE, row.names = TRUE, append = FALSE)

saveRDS(gl_ref_sites_df,paste(data_folder, paste('/TimedSwimSites_NEGreenlip_ReferenceSites_Final_2025.RDS', sep = ''), sep = ''))

```

```{r site selection post 2024}

# Specify assessment year
assess_year <- 2025

# Create buffer around reference sites to exclude new sites within 50 m for proceeding year site selections.
gl_ref_sites_buff <- st_buffer(gl_ref_sites, dist = 50) %>% 
 st_transform(crs = st_crs(4326))

# Remove sites within 50 m of reference sites
gl_ts_hex_sf <- ts.hex %>%
 st_centroid() %>% 
 st_transform(crs = st_crs(4326)) %>% 
 erase.point(., gl_ref_sites_buff)

# Create vector of all OIDs sampled to filter from next assessment year random site selection
oid_samp <- gl_sites %>% 
        distinct(site, .keep_all = T) %>% 
        filter(!is.na(oid)) %>% 
        ungroup() %>% 
        pull(oid)

# Sites to exclude (oid) where a gps was left on while transiting on a motherboat.
oid_exc <- c(1062044, 1062045, 1062604, 1062605, 1062606, 1062607, 1062608, 1062609, 1062610, 1062611, 1062614, 1062632, 1062633, 1062634, 1062635, 1063208, 1063209, 1063210, 1063211, 1063212, 1063213, 1063214, 1063215, 1063216, 1063217, 1063218, 1063219, 1063220, 1063221, 1063222, 1063223, 1063224, 1063225, 1063226, 1063227, 1063228)

set.seed(123) 

# set sample size required (i.e. n = 45)
samp.size <- 45

# randomly select sites (NOTE: manually change dataframe name for each block)
cellqt_NE_GL <- gl_ts_hex_sf %>% 
  filter(!oid %in% c(oid_exc, oid_samp)  &
           cell.ntile > 2) %>%
  group_by(blockno, cell.ntile)  %>% 
  st_as_sf() %>%
  subsample.distance(size = samp.size, d = 50) %>% #50 m distance between sites
  st_as_sf() %>% 
  ungroup()

  # quick summary of sites x block x strata
table(cellqt_NE_GL$blockno, cellqt_NE_GL$cell.ntile)

# Find the centroid of each hex cell to create point/site location coordinates
ts_sites_pointqt_sf <- st_centroid(cellqt_NE_GL)

# Convert geometry to latitude and longitude to create dataframe
ts_sites_df <- ts_sites_pointqt_sf %>%
        st_transform(crs = st_crs(4326)) %>%
  sfheaders::sf_to_df(fill = T) %>%
  select(c(zone, blockno, subblockno, cell.ntile, oid, x, y)) %>%
  dplyr::rename('latitude' = y,
                'longitude' = x,
                'cell_ntile' = cell.ntile)

# Add site names using naming convention 'AB-sampyear-blockno-site.order'
ts_sites_df <- ts_sites_df %>% 
 group_by(blockno) %>% 
  mutate(site_no = row_number(),
         site = paste('GL', assess_year, blockno, site_no, sep = '-')) %>% 
 ungroup()

# Denote as new sites
ts_sites_df <- ts_sites_df %>% 
 mutate(ref_site = 'new')

# Unset seed
set.seed(NULL)

# Select data 
ts_sites_df <- ts_sites_df %>% 
  select(site, blockno, subblockno, oid, cell_ntile, longitude, latitude, ref_site)

# Join new sites with original reference sites
ts_sites_final_df <- bind_rows(gl_ref_sites_df, ts_sites_df) %>% 
 arrange(blockno)

# Convert df to SF for export
ts_sites_final_sf  <-  ts_sites_final_df %>% 
  st_as_sf(coords = c('longitude', 'latitude'), crs = st_crs(4326))

# Create GPX file for export
ts_sites_final_gpx <- ts_sites_final_df %>% 
 dplyr::rename('name' = 'site') %>% 
 select(c(name, longitude, latitude)) %>% 
 add_column(description = NA, 
            comment = NA)

```

# Export data
```{r export data}

# Export folder path

gis_folder_path <- paste(file.path(paste(sprintf('C:/Users/%s/Dropbox (UTAS Research)/DiveFisheries/GIS/SpatialLayers/TimeSwimLayers',
Sys.info()[["user"]]), sep = '')))

data_folder_path <- paste(file.path(paste(sprintf('C:/Users/%s/Dropbox (UTAS Research)/DiveFisheries/Abalone/FISdata/', Sys.info()[["user"]]), paste('FIS_TimedSwimSurveys', assess_year, sep = ''), sep = '')))

# Export geospatial package for GIS and KML for Google Earth

outname_point <- paste(gis_folder_path, paste('/TimedSwimSites_NEGreenlip_Final_', assess_year, ".gpkg", sep = ""), sep = '')
outname_point <- paste(data_folder_path, paste('/TimedSwimSites_NEGreenlip_Final_', assess_year, ".gpkg", sep = ""), sep = '')

st_write(ts_sites_final_sf, dsn = outname_point, layer = "pointqt", driver = "GPKG", append = F)

# Export KML for Google Earth

outname_point_kml <- paste(gis_folder_path, paste('/TimedSwimSites_NEGreenlip_Final_', assess_year, ".kml", sep = ""), sep = '')
outname_point_kml <- paste(data_folder_path, paste('/TimedSwimSites_NEGreenlip_Final_', assess_year, ".kml", sep = ""), sep = '')

st_write(ts_sites_final_sf, dsn = outname_point_kml, layer = "pointqt", driver = "kml", append = F) 
 
# Export Excel files 
write.xlsx(ts_sites_final_df, paste(gis_folder_path, paste('/TimedSwimSites_NEGreenlip_Final_', assess_year, ".xlsx", sep = ""), sep = ''), sheetName = "Sheet1", colNames = TRUE, rowNames = TRUE, append = FALSE)
write.xlsx(ts_sites_final_df, paste(data_folder_path, paste('/TimedSwimSites_NEGreenlip_Final_', assess_year, ".xlsx", sep = ""), sep = ''), sheetName = "Sheet1", colNames = TRUE, rowNames = TRUE, append = FALSE)


# Export Excel file for GPX conversion
write.xlsx(ts_sites_final_gpx, paste(gis_folder_path, paste('/TimedSwimSites_NEGreenlip_Final_', assess_year, '_GPX', ".xlsx", sep = ""), sep = ''), sheetName = "Sheet1", colNames = TRUE, rowNames = TRUE, append = FALSE)
write.xlsx(ts_sites_final_gpx, paste(data_folder_path, paste('/TimedSwimSites_NEGreenlip_Final_', assess_year, '_GPX', ".xlsx", sep = ""), sep = ''), sheetName = "Sheet1", colNames = TRUE, rowNames = TRUE, append = FALSE)
  
```