---
title: "Selecting locations for NE Greenlip Timed-swim surveys"
author: "Jaime McAllister and Craig Mundy"
date: "`r Sys.Date()`"
output:
  pdf_document:
    fig_height: 8
    fig_width: 8
    # includes:
    #   in_header: header.tex
    toc: yes
  word_document:
    fig_caption: yes
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = "",     ## nothing goes in front of each line
                      message = FALSE,  ## into the console, not the document
                      warning = FALSE,  ## into the console, not the document
                      fig.align = "center",
                      fig.show = "hold",
                      out.height = "0.8\\textwidth",
                      out.width = "0.8\\textwidth")
options(knitr.kable.NA = '')


suppressPackageStartupMessages({
library(tictoc)
library(sf)
library(nngeo)
library(maptools)
library(rgdal)
library(sp)
library(spatialEco)  
library(rgeos)
library(lubridate)
library(spdplyr)
library(tidyverse)
library(tmap)
library(tictoc)
library(spdep)
library(openxlsx)  
})

## set EPSG CRS

GDA2020 <- st_crs(7855)
GDA94 <- st_crs(28355)
WGS84 <- st_crs(4326)


```


# Read in hex layer
Read in grid wide spatial layer, and transform to GDA2020. This hex cell layer contains the majority of fishing activity in the closed blocks.
```{r prepare base layers}
## Read in greenlip abalone wide hex layer

ab.hex <- readRDS(sprintf("C:/Users/%s/University of Tasmania/IMAS-DiveFisheries - Assessments - Documents/Assessments/GIS/SpatialLayers/TimeSwimLayers/grid1HA_greenlip.rds",
        Sys.info()[["user"]])) 


## convert to sf
sf.ab.hex <- st_as_sf(ab.hex)

## Transform from GDA94 to GDA2020
sf.ab.hex <- st_transform(sf.ab.hex, GDA2020)
```

# Add quartile to dataframe
THe input variable is the total catch in Kg for each cell from 2012 - 2019. We use the dplyr ntile() function to calculate five quantiles (0-20, 20-40, etc). based on total catch.
```{r add quartile var}

# Filter hex layer to required block and add quartile
ts.hex <- sf.ab.hex %>% 
  mutate(subblockno = gsub(" ", "", subblockno)) %>% 
  filter(subblockno %in%  c('31A', '31B', '39A','39B')) %>% 
  within({
    cell.ntile <- ntile(GLhexcatch, 5)
  })

outname.ts.hex <- sprintf("C:/Users/%s/University of Tasmania/IMAS-DiveFisheries - Assessments - Documents/Assessments/GIS/SpatialLayers/grid1HA_greenlip.gpkg", Sys.info()[["user"]])
st_write(ts.hex, dsn = outname.ts.hex, layer = "pointqt", driver = "GPKG", append = F)

hex.cell.summary <- ts.hex %>%
  st_set_geometry(NULL) %>%
  group_by(cell.ntile) %>%
  summarise(
    mncatch = mean(GLhexcatch, na.rm = T),
    catch = sum(GLhexcatch, na.rm = T),
    n = n()
  ) %>%
  print()

```

# randomly sample cells within strata.
We want to randomly select 150 hex cells across the top three quantile strata. The two lower quantile strata contribute `r sum(hex.cell.summary$catch[1:2])/sum(hex.cell.summary$catch)`  of a total catch of `r sum(hex.cell.summary$catch) `. The top three quantile groups contribute `r sum(hex.cell.summary$catch[3])/sum(hex.cell.summary$catch)`, `r sum(hex.cell.summary$catch[4])/sum(hex.cell.summary$catch)`  and `r sum(hex.cell.summary$catch[5])/sum(hex.cell.summary$catch)`  of the catch, respectively.
```{r random sample of cells}

set.seed(123) 

# set sample size required (i.e. n = 150)
samp.size <- 150

# Exclude oid where a gps was left on while transiting on a motherboat.

oid_exc <- c(1062044, 1062045, 1062604, 1062605, 1062606, 1062607, 1062608, 1062609, 1062610, 1062611, 1062614, 1062632, 1062633, 1062634, 1062635, 1063208, 1063209, 1063210, 1063211, 1063212, 1063213, 1063214, 1063215, 1063216, 1063217, 1063218, 1063219, 1063220, 1063221, 1063222, 1063223, 1063224, 1063225, 1063226, 1063227, 1063228)

# randomly select sites (NOTE: manually change dataframe name for each block)
cellqt_NE_GL <- ts.hex %>% 
  filter(!oid %in% oid_exc &
           cell.ntile > 2) %>%
  group_by(blockno, cell.ntile)  %>% 
  st_as_sf() %>%
  subsample.distance(size = samp.size, d = 50) %>% 
  st_as_sf() %>% 
  ungroup()

  # quick summary of sites x block x strata
table(cellqt_NE_GL$blockno, cellqt_NE_GL$cell.ntile)

```
# Examine Nearest Neighbour distances
We find the  k=1 Nearest Neighbour and then calculate the distance between a cell and its closest neighbor. Centroids of hex cells are ~ 107m apart.
```{r examine nnb}
knear <- knearneigh(st_centroid(cellqt_NE_GL), k = 1, longlat = FALSE, use_kd_tree = TRUE)
neighbors <- knn2nb(knear)

summary.nb(neighbors, st_coordinates(st_centroid(cellqt_NE_GL)), longlat = NULL, zero.policy=TRUE)

coords.hex <- st_coordinates(st_centroid(cellqt_NE_GL))
nc_sp <- as(cellqt_NE_GL, 'Spatial')
cellqt_NE_GL.Knbdist <- nbdists(neighbors, st_coordinates(st_centroid(cellqt_NE_GL)), longlat = NULL)

nbdists <- unlist(cellqt_NE_GL.Knbdist) %>% as.tibble()

nbdists %>% filter(nbdists < 125) %>% count()

colnames(nbdists) <- "nbDist"

nbdists %>% 
  ggplot(aes(nbDist)) +
  geom_histogram()

nbdists %>% 
  filter(nbDist < 2000) %>% 
  ggplot(aes(nbDist)) +
  geom_histogram()
```

# Plot of nnb connections
```{r examine nnb network}

neighbors_sf <- as(nb2lines(neighbors, coords = st_coordinates(st_centroid(cellqt_NE_GL))), 'sf' )
neighbors_sf <- st_set_crs(neighbors_sf, st_crs(cellqt_NE_GL))

ggplot(cellqt_NE_GL) + 
  geom_sf(fill = 'salmon', color = 'white') +
  geom_sf(data = neighbors_sf) +
  geom_sf(data = st_centroid(cellqt_NE_GL)) +
  theme_minimal() +
  ylab("Latitude") +
  xlab("Longitude")



```


# Export files
It is much easier to explore the data in QGIS. We export the selected cells as a hex layer and a point layer (centroid of the hex).

```{r export to gpkg}

## Export cells to gpkg ####
pointqt <- st_centroid(cellqt_NE_GL)
outname.point <- sprintf("C:/Users/%s/University of Tasmania/IMAS-DiveFisheries - Assessments - Documents/Assessments/GIS/SpatialLayers/TimeSwimLayers/TimedSwim_NE_GL_2023_50m.gpkg", Sys.info()[["user"]])
st_write(pointqt, dsn = outname.point, layer = "pointqt", driver = "GPKG", append = F)

# Convert geometry to latitude and longitude to create dataframe
pointqt_df <- pointqt %>%
        st_transform(crs = st_crs(4326)) %>%
  sfheaders::sf_to_df(fill = T) %>%
  select(c(zone, blockno, subblockno, cell.ntile, oid, x, y)) %>%
  dplyr::rename('latitude' = y,
                'longitude' = x) %>% 
 arrange((oid)) %>% 
 mutate(site_order = row_number(),
        site = paste('GL', 2023, blockno, site_order, sep = '-'))

pointqt_sf  <-  pointqt_df %>% 
  st_as_sf(coords = c('longitude', 'latitude'), crs = st_crs(4326))
 
# Export Excel file 
write.xlsx(pointqt_df, sprintf("C:/Users/%s/University of Tasmania/IMAS-DiveFisheries - Assessments - Documents/Assessments/GIS/SpatialLayers/TimeSwimLayers/TimedSwim_NE_GL_2023_50m.xlsx", Sys.info()[["user"]]), 
           sheetName = "Sheet1",
           col.names = TRUE, row.names = TRUE, append = FALSE)
 
# Export Excel file ready for GPX external file creation for plotter
# GDAL package for creating GPX files no longer supported in R
# pointqt_gpx <- pointqt_df %>% 
#    select(site, longitude, latitude) %>% 
#    rename('waypointid' = site)

pointqt_gpx <- pointqt_df %>% 
 dplyr::rename('name' = 'site') %>% 
 select(c(name, longitude, latitude)) %>% 
 add_column(description = NA, 
            comment = NA)

write.xlsx(pointqt_gpx, sprintf("C:/Users/%s/University of Tasmania/IMAS-DiveFisheries - Assessments - Documents/Assessments/GIS/SpatialLayers/TimeSwimLayers/TimedSwim_NE_GL_2023_GPX.xlsx", Sys.info()[["user"]]), 
           sheetName = "Sheet1",
           col.names = TRUE, row.names = F, append = FALSE)
 
```
