---
title: "Tasmanian Seafoods Commercial Catch Sampling"
author:
  - name: Jaime McAllister
    affiliations:
        - name: IMAS, University of Tasmania
          department: IMAS-FA
date: last-modified
date-format: "[Last Updated on] DD MMMM, YYYY"
format:
  docx:
    highlight-style: github
    papersize: A4
    code-overflow: "wrap"
    reference-doc: word-styles-reference-01.docx
    toc: true
    number-sections: false
    toc-depth: 4
    number-depth: 4
    margin-left: 0.75in
    margin-right: 0.75in
    margin-top: 1in
    margin-bottom: 1in
  pdf:
    documentclass: scrreport
    keep-tex:  true
    dpi: 600
    pdf-engine: lualatex
    toc: true
    toc-depth: 4
    toc_float: true
    number-sections: false
    number-depth: 4
    highlight-style: github
    papersize: "A4paper"
    linestretch: 1.25
    mainfont: Calibri
    geometry:
      - left = 20mm
      - right = 20mm
      - top = 20mm
      - bottom = 10mm
editor: 
  markdown: 
    wrap: 72
---
\newpage

```{r setup}
#| echo: false
#| warning: false
#| message: false

##---------------------------------------------------------------------------##
# Clear console
rm(list = ls())

##---------------------------------------------------------------------------##
## 1. Load libraries ####
suppressPackageStartupMessages({
        library(dplyr)
        library(ggplot2)
        library(scales)
        library(tidyr)
        library(gdata)
        library(openxlsx)
        library(lubridate)
        library(reshape)
        library(gridExtra)
        library(ggpubr)
        library(readxl)
        library(tibble)
        library(data.table)
        library(janitor)
        library(anytime)
        library(stringr)
        library(broom)
        library(purrr)
        library(sf)
        library(ggspatial)
        library(tmap)
        library(sf)
        library(sp)
        library(RColorBrewer)
        library(viridis)
        library(ggpmisc)
        library(arsenal)
  library(fuzzyjoin)
 library(tidytext)
})

source("C:/GitCode/AbResearch/getLegend.r")
source("C:/GitCode/AbResearch/StandardError_Functions.r")

```


```{r load data}
#| echo: false
#| warning: false
#| message: false


##-------------------------------------------------------------------------------------------------------##
# Load data ####

# Load most recent compiled MM dataframe

compiledMM_df_final <- readRDS(paste(sprintf('C:/Users/%s/Dropbox (UTAS Research)/DiveFisheries/Abalone/MMdata/compiledMM.df.final.RDS',
                                            Sys.info()[["user"]])))


# Identify stock assessment year of interest
target_year <- 2025


# Create plots folder
mm_plots_folder <- file.path(paste(sprintf('C:/Users/%s/Dropbox (UTAS Research)/DiveFisheries/Abalone/Assessment/Figures/MM',
                                           Sys.info()[["user"]])),
                             paste(target_year, '/', 'MM_Plots_', target_year, sep = ''))


```

# Weight Grading

Since 2020, catches have been predominantly composed of medium grade abalone (@fig-stacked). However, the proportion of large grade abalone has steadily increased, particularly from 2022 onward, while small grade abalone have declined in relative proportion. 

```{r stacked grade plot}
#| label: fig-stacked
#| fig-cap: "Stacked frequency distribution of weight gradings across extra small, small, medium and large abalone, illustrating the relative contribution of each weight grade to the overall distributions in the Western Zone between 2020 and 2025. Number of abalone measured above each boxplot."
#| fig-pos: "H"
#| fig-width: 10
#| fig-height: 8
#| out-height: 90%
#| echo: false
#| warning: false
#| message: false

# Stacked barplot of grades measured for zone in stock assessment year
    
# Determine number of abalone measured by grade per zone

  blockno.grade.meas <- compiledMM_df_final %>% 
    filter(!is.na(whole.weight) &
             between(fishyear, target_year - 5, fishyear) &
             numblocks == 1) %>% 
    group_by(fishyear, newzone, grade) %>% 
    summarise(grade.meas = n()) %>% 
    mutate(zone.blockno = paste(newzone, sep = '-'))

# Create plot label for total number of abalone measured per block
  blockno_n_plot_lab <- compiledMM_df_final %>% 
   filter(!is.na(whole.weight) &
           between(fishyear, target_year - 5, fishyear) &
           numblocks == 1) %>% 
    group_by(fishyear, newzone) %>% 
    summarise(n = n()) %>% 
    mutate(zone.blockno = paste(newzone, sep = '-'),
           y.pos = 1.02)

# Create 100% stacked barplot of grades for western zone
  
  grade_stack_plot <- ggplot(blockno.grade.meas %>% filter(newzone == 'W'), 
                             aes(x = fishyear, 
                                 y = grade.meas,
                             fill = forcats::fct_rev(grade))) + 
    geom_bar(position = "fill", stat = "identity") + 
    scale_y_continuous(labels = scales::percent_format())+
    xlab('Year')+
    ylab('Percentage')+
    geom_text(data = blockno_n_plot_lab %>% filter(newzone == 'W') , aes(x = fishyear, y = y.pos, label = n, fill = NULL), size = 3)+
    scale_fill_manual(name = 'Grade', 
                      labels = c('XSmall', 'Small', 'Medium', 'Large'),
                      values = c('#868686FF', '#EFC000FF', '#0073C2FF','#CD534CFF'))+
    theme_bw()
  
  grade_stack_plot
    
# Save plot
  
  ggsave(
    filename = paste(mm_plots_folder, 'MM_ZoneNoGradeSummary_W_', target_year, '.pdf', sep = ''),
    plot = grade_stack_plot,
    width = 7.4,
    height = 5.57,
    units = 'in'
  )
  ggsave(
    filename = paste(mm_plots_folder, 'MM_ZoneNoGradeSummary_W_', target_year, '.wmf', sep = ''),
    plot = grade_stack_plot,
    width = 7.4,
    height = 5.57,
    units = 'in'
  )
  ggsave(
    filename = paste(mm_plots_folder, 'MM_ZoneNoGradeSummary_W_', target_year, '.png', sep = ''),
    plot = grade_stack_plot,
    width = 7.4,
    height = 5.57,
    units = 'in'
  )

```

# Length Frequency

Size composition in the western zone has remained relatively stable since 2020, with catches consistently centered around a median length of 155–160 mm (@fig-boxplot). In 2025, catches have been dominated by generally larger abalone, with a median length closer to 160 mm.

```{r median length}
#| label: fig-boxplot
#| fig-cap: 'Boxplot of length frequency distributions between 2020 and 2025 for the Western Zone. Red line under each boxplot indicates the LML for that year. Number of abalone measured given above each boxplot.'
#| fig-pos: "H"
#| fig-width: 10
#| fig-height: 8
#| out-height: 90%
#| echo: false
#| warning: false
#| message: false

# Boxplot of lengths measured for zone in stock assessment year
    
plotdat_zone <- compiledMM_df_final %>%
filter(!is.na(whole.weight) &
             between(fishyear, target_year - 5, fishyear) &
             numblocks == 1 &
        newzone == 'W')

 # generate a count of records for each year to add to boxplot
      
      plotdat_n <- plotdat_zone %>%
        group_by(fishyear, newzone) %>%
        summarize(n = n())
      
      # generate table of size limits for each year to add to boxplot
      
      size_limit <- plotdat_zone %>% 
        group_by(fishyear, newzone) %>% 
        summarise(legal.min.length = max(sizelimit))
      
      # generate boxplot of shell lengths for chosen grouping variable
      
      mm_zone_boxplot <-
        ggplot(plotdat_zone %>% filter(newzone == 'W'), aes(x = fishyear, y = shell.length, group = fishyear)) +
        geom_boxplot(outlier.colour = "orange", outlier.size = 1.5) +
        geom_text(
          data = plotdat_n %>% filter(newzone == 'W'),
          aes(y = 200, label = n),
          size = 3,
          angle = 0
        ) +
        geom_point(data = size_limit %>% filter(newzone == 'W'), aes(x = fishyear, y = legal.min.length), 
                   shape = 95, size = 7, colour = "red")+
        xlab('Year') +
        ylab(paste('Shell Length (mm)')) +
        coord_cartesian(ylim = c(120, 200)) +
        theme_bw() +
        theme(
          plot.title = element_text(hjust = 0.5),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          axis.line = element_line(colour = "black"),
          axis.text.x = element_text(angle = 0, vjust = 0.5)
        )
      
mm_zone_boxplot

  ggsave(
    filename = paste(mm_plots_folder, 'MM_ZoneSizeStructure_W_', target_year, '.pdf', sep = ''),
    plot = mm_zone_boxplot,
    width = 7.4,
    height = 5.57,
    units = 'in'
  )
  ggsave(
    filename = paste(mm_plots_folder, 'MM_ZoneSizeStructure_W_', target_year, '.wmf', sep = ''),
    plot = mm_zone_boxplot,
    width = 7.4,
    height = 5.57,
    units = 'in'
  )
  ggsave(
    filename = paste(mm_plots_folder, 'MM_ZoneSizeStructure_W_', target_year, '.png', sep = ''),
    plot = mm_zone_boxplot,
    width = 7.4,
    height = 5.57,
    units = 'in'
  )
      
```    

